---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SynSigGen

<!-- badges: start -->

[![R-CMD-check-v1](https://github.com/steverozen/SynSigGen/workflows/R-CMD-check-v1/badge.svg?branch=1.1.1-branch)](https://github.com/steverozen/SynSigGen/actions?query=workflow%3AR-CMD-check-v1+branch%3A1.1.1-branch)
[![R-CMD-check-bioc-3.14](https://github.com/steverozen/SynSigGen/workflows/R-CMD-check-bioc-3.14/badge.svg?branch=1.1.1-branch)](https://github.com/steverozen/SynSigGen/actions?query=workflow%3AR-CMD-check-bioc-3.14+branch%3A1.1.1-branch)
<!-- badges: end -->

Synthetic (Mutational) Signature Generation (**`SynSigGen`**)

## Purpose

Create catalogs of synthetic mutational spectra for assessing the
performance of mutational signature analysis programs. 'SynSigGen' stands
for Generation of Synthetic Signatures and Spectra.

<br>

## Installation

Before installation, prerequisites in 
[Bioconductor](https://www.bioconductor.org/)
needs to be installed:

```r
install.packages("BiocManager")
BiocManager::install(
  c("Biostrings", "BSgenome", "GenomeInfoDb", "GenomicRanges")
)
```

Install from GitHub with the R command line:

``` r
install.packages("remotes")
remotes::install_github(repo = "steverozen/SynSigGen", ref = "1.2.1-branch")
```

<br>

## Example usage

``` {r include = FALSE}
require(SynSigGen)
```

### 1. Generate synthetic tumor spectra used in *The repertoire of mutational signatures in human cancer*, Alexandrov et al. (2020)

Use functions below to generate 11 spectra datasets used in paper
*The repertoire of mutational signatures in human cancer* 
(https://doi.org/10.1038/s41586-020-1943-3), published in _Nature_.
The data sets are available at 
[Synapse](https://www.synapse.org/#!Synapse:syn18497223):

``` r
# Users should specify regress.dir = NULL unless for comparison
# with original data sets.
# 
# Compare tools (e.g., BeyondCompare, Meld) are recommended 
# over specifying regress.dir,
# because the latter might raise an error even when query 
# and original data sets are identical.
#
# Users should specify top.level.dir to the destination folder
# for data sets. Otherwise default paths will be used.
PancAdenoCA1000()
ManyTypes2700()
RCCOvary1000()
Create.3.5.40.Abstract()
BladderSkin1000()
Create.2.7a.7b.Abstract()
CreateRandomSyn()
```

The description of 11 data sets are available at section
*"Description of each suite of synthetic data sets"*
in Supplementary Note 2 of the paper.

### 2. Generate synthetic spectra used in paper *Accuracy of Mutational Signature Software on Correlated Signatures*, Wu et al. (2022)

To generate 20 spectra data sets with mutation load of 
and correlation between SBS1 and SBS5 varied, use function 

``` r
CreateSBS1SBS5CorrelatedSyntheticData()
```

This paper is published at [_Scientific Reports_](https://doi.org/10.1038/s41598-021-04207-6),
and the original data sets are available at 
[Zenodo](https://doi.org/10.5281/zenodo.2636980).

### 3. Generate synthetic spectra used in paper *mSigHdp: hierarchical Dirichlet processes for mutational signature discovery*, Liu et al. (2022)

To generate 3 spectra data sets on single base substitution (SBS)
mutation channels and 3 spectra datasets on indel channels,
check GitHub repository Liu_et_al_Sup_Files. The dataset generation codes
in this repository requires SynSigGen >= 1.1.1 as a dependency.

<br>

## Notes for functions to generate legacy data sets (1 & 2)

* The wrapper functions used to generate data sets in [_Nature_ paper](https://doi.org/10.1038/s41586-020-1943-3)
  are in R files with suffix "_Nat".

* The wrapper functions used to generate data sets for paper on 20 
  correlated data sets are in file “CreateSynSBS1SBS5Correlated.R”.

* These wrapper functions are primarily used to generate legacy data sets,
  as they don't round the exposures to integers. 

* By contrast, `GenerateSyntheticExposures()`
  now rounds the exposures by default from version 1.0.10.

<br>

## Reference manual
https://github.com/steverozen/SynSigGen/blob/1.2.1-branch/data-raw/SynSigGen_1.2.1.pdf
