---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SynSigGen

<!-- badges: start -->

  [![Travis build status](https://travis-ci.com/steverozen/SynSigGen.svg?branch=master)](https://travis-ci.com/steverozen/SynSigGen)

<!-- badges: end -->

Synthetic (Mutational) Signature Generation ('SynSigGen')

## Purpose

Create catalogs of synthetic mutational spectra for assessing the
  performance of mutational signature analysis programs. 'SynSigGen' stands
  for Generation of Synthetic Signatures and Spectra.

## Installation

Before installation, prerequisites in 
[Bioconductor](https://www.bioconductor.org/)
needs to be installed:

```r
install.packages("BiocManager")
BiocManager::install(c("Biostrings","BSgenome","GenomeInfoDb","GenomicRanges"))
```

Install from GitHub with the R command line:

``` r
install.packages("devtools")
devtools::install_github("steverozen/SynSigGen")
```

## Example usage

``` {r include = FALSE}
require(SynSigGen)
```

### Generate synthetic tumor spectra used in paper *The repertoire of mutational signatures in human cancer*

Use functions below to generate 11 spectra datasets used in paper
*The repertoire of mutational signatures in human cancer* 
(https://doi.org/10.1038/s41586-020-1943-3):

``` r
PancAdenoCA1000()
ManyTypes2700()
RCCOvary1000()
Create.3.4.40.Abstract()
BladderSkin1000()
Create.2.7a.7b.Abstract()
CreateRandomSyn()
```

### Generate synthetic spectra used in paper *"Performance of Mutational Signature Software on Correlated Signatures"*

To generate 20 spectra data sets with mutation load of 
and correlation between SBS1 and SBS5 varied, use function 

``` r
CreateSBS1SBS5CorrelatedSyntheticData()
```



## Reference manual
https://github.com/steverozen/SynSigGen/blob/master/data-raw/SynSigGen_1.0.4.pdf
