---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SynSigGen

<!-- badges: start -->

[![R-CMD-check](https://github.com/steverozen/SynSigGen/workflows/R-CMD-check/badge.svg?branch=1.1.0-branch)](https://github.com/steverozen/SynSigGen/actions?query=workflow%3AR-CMD-check+branch%3A1.1.0-branch)


[![R-CMD-check-bioc](https://github.com/steverozen/SynSigGen/workflows/R-CMD-check-bioc/badge.svg?branch=1.1.0-branch)](https://github.com/steverozen/SynSigGen/actions?query=workflow%3AR-CMD-check-bioc+branch%3A1.1.0-branch)


<!-- badges: end -->

Synthetic (Mutational) Signature Generation (**`SynSigGen`**)

## Purpose

Create catalogs of synthetic mutational spectra for assessing the
  performance of mutational signature analysis programs. 'SynSigGen' stands
  for Generation of Synthetic Signatures and Spectra.

## Installation

Before installation, prerequisites in 
[Bioconductor](https://www.bioconductor.org/)
needs to be installed:

```r
install.packages("BiocManager")
BiocManager::install(c("Biostrings","BSgenome","GenomeInfoDb","GenomicRanges"))
```

Install from GitHub with the R command line:

``` r
install.packages("remotes")
remotes::install_github(repo = "steverozen/SynSigGen", ref = "master")
```

## Example usage

``` {r include = FALSE}
require(SynSigGen)
```

### 1. Generate synthetic tumor spectra used in paper *The repertoire of mutational signatures in human cancer*

Use functions below to generate 11 spectra datasets used in paper
*The repertoire of mutational signatures in human cancer* 
(https://doi.org/10.1038/s41586-020-1943-3), published in _Nature_.
The data sets are available at [Synapse](https://www.synapse.org/#!Synapse:syn18497223):

``` r
# Users should specify regress.dir = NULL unless for comparison
# with original data sets.
# 
# Compare tools (e.g. BeyondCompare, Meld) is recommended 
# over specifying regress.dir,
# because the latter might raise an error even when query 
# and original data sets are identical.
#
# Users should specify top.level.dir to the destination folder
# for data sets. Otherwise default paths will be used.
PancAdenoCA1000()
ManyTypes2700()
RCCOvary1000()
Create.3.5.40.Abstract()
BladderSkin1000()
Create.2.7a.7b.Abstract()
CreateRandomSyn()
```

The description of 11 data sets are available at section *"Description of each suite of synthetic data sets"* in Supplementary Note 2 of the paper. 

### Generate synthetic spectra used in paper *"Accuracy of Mutational Signature Software on Correlated Signatures"*

To generate 20 spectra data sets with mutation load of 
and correlation between SBS1 and SBS5 varied, use function 

``` r
CreateSBS1SBS5CorrelatedSyntheticData()
```

The original data sets are available at [Zenodo](https://doi.org/10.5281/zenodo.2636980).

## Notes for functions to generate legacy data sets (1 & 2)

* The wrapper functions used to generate data sets in [_Nature_ paper](https://doi.org/10.1038/s41586-020-1943-3) are in R files 
with suffix "_Nat".

* The wrapper functions used to generate data sets for paper on 20 
correlated data sets are in file “CreateSynSBS1SBS5Correlated.R”.

* These wrapper functions are primarily used to generate legacy data sets,
as they don't round the exposures to integers. 

* By contrast, `GenerateSyntheticExposures()`
now rounds the exposures by default from version 1.0.10.


## Reference manual
https://github.com/steverozen/SynSigGen/blob/1.1.0-branch/data-raw/SynSigGen_1.1.0.pdf
